---
title: "02 - The Eaton and Kortum model"
author: "Rafael F. Bressan"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    css: "style.css"
bibliography: "references.bib"
link-citations: yes
---

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: { 
      equationNumbers: { 
            autoNumber: "all",
            formatNumber: function (n) {return n}
      } 
  }
});
</script>

## Theoretical Framework

In this section we will briefly describe the theoretical model as in @costinot2012goods, which is based on @eaton2002technology. It consists of an economy with multiple-countries, multiple industries, and labor as the solely production factor. Intra-industry heterogeneity is allowed, as in @eaton2002technology, each good is available in many varieties.

Consider a world economy comprising $i=1, \ldots, I$ countries and labor as the only factor of production. There are $k=1, \ldots, K$ industries or goods. Labor is perfectly mobile across industries and immobile across countries. We denote by $L_{i}$ and $w_{i}$ the number of workers and the wage in country $i$, respectively. 

**Technology.** Each good $k$ may come in an infinite number of varieties indexed by $\omega \in \Omega$ We denote by $z_{i}^{k}(\omega)$ the number of units of the $\omega$ th variety of good $k$ that can be produced with one unit of labor in country $i$. Assume the following:

**A1.** For all countries $i$, goods $k$, and their varieties $\omega$,  $z_{i}^{k}(\omega)$ is a random variable drawn independently for each triplet $(i, k, \omega)$ from a Fréchet distribution $F_{i}^{k}(\cdot)$ such that
$$
F_{i}^{k}(z)=\exp \left[-\left(z / z_{i}^{k}\right)^{-\theta}\right], \quad \text { for all } z \geq 0
$$
where $z_{i}^{k}>0$ and $\theta>1$.
Technological differences across countries and industries only depend on **two parameters**, $z_{i}^{k}$ and $\theta$. We refer to $z_{i}^{k}$ as the _fundamental productivity_ of country $i$ in industry $k$ while $\theta$ is the _intra-industry heterogeneity_.

**A2.** For each unit of good $k$ shipped from country $i$ to country $j$, only $1 / d_{i j}^{k} \leq 1$ units arrive, with $d_{i j}^{k}$ such that 

$$
\begin{cases}
d_{i i}^{k}=1 \text{, and}\\ 
d_{i l}^{k} \leq d_{i j}^{k} \cdot d_{j l}^{k} \text{, for any third country } l.
\end{cases}
$$

**Market structure.** Markets are assumed to be perfectly competitive. Together with constant returns to scale in production, perfect competition implies the following:

**A3.** In any country $j$, the price $p_{j}^{k}(\omega)$ paid by buyers of variety $\omega$ of good $k$ is
$$
p_{j}^{k}(\omega)=\min _{1 \leq i \leq I}\left[c_{i j}^{k}(\omega)\right]
$$
where $c_{i j}^{k}(\omega)=\left(d_{i j}^{k} \cdot w_{i}\right) / z_{i}^{k}(\omega)$ is the cost of producing and delivering one unit of this variety from country $i$ to country $j$.

Buyers in country $j$ are "shopping around the world" for the best price available. In what follows, we let $c_{i j}^{k}=\left(d_{i j}^{k} \cdot w_{i}\right) / z_{i}^{k}>0$.

**Preferences.** The upper tier utility function is Cobb-Douglas, while the lower tier is CES[^ds]. Accordingly, expenditures are such that:

**A4.** In any country $j$, total expenditure on variety $\omega$ of good $k$ is
$$
x_{j}^{k}(\omega)=\left[p_{j}^{k}(\omega) / p_{j}^{k}\right]^{1-\sigma_{j}^{k}} \cdot \alpha_{j}^{k} w_{j} L_{j}
$$

[^ds]: Demand comes from the usual Dixit-Stiglitz aggregator model. The basics of this model can be seen [here](http://www.columbia.edu/~jid2106/td/dixitstiglitzbasics.pdf).

where $p_j^k\equiv[\sum_{\omega'\in\Omega} p_j^k(\omega)^{1-\sigma_j^k}]^{1/(1-\sigma_j^k)}$ is the price index _within_ industry $k$ of country $j$. The country $j$'s consumer price index is defined as: $p_j\equiv\Pi_{k=1}^K(p_j^k)^{\alpha_j^k}$.

**Trade balance.** Denote by $x_{i j}^{k} \equiv \sum_{\omega \in \Omega_{i j}^{k}} x_{j}^{k}(\omega)$ the value of total exports from country $i$ to country $j$ in industry $k$, where $\Omega_{i j}^{k}$ is the set of varieties exported by country $i$ to country $j$ in industry $k$. Similarly, denote by $\pi_{i j}^{k} \equiv x_{i j}^{k} / \sum_{i^{\prime}=1}^{I} x_{i^{\prime} j}^{k}$ the share of exports from country $i$ in country $j$ and industry $k$. The final assumption is the following:

**A5.** For any country $i$, trade is balanced
$$
\sum_{j=1}^{I} \sum_{k=1}^{K} \pi_{i j}^{k} \alpha_{j}^{k} \gamma_{j}=\gamma_{i},
$$
where $\gamma_{i} \equiv w_{i} L_{i} / \sum_{i^{\prime}=1}^{I} w_{i^{\prime}} L_{i^{\prime}}$ is the share of country $i$ in world income.

## Calibration

We start the calibration by choosing an appropriate value for $\theta$. From Table 3 in @costinot2012goods I selected column (3), $\theta = 6.534$, where the method of estimation was an IV, to be used throughout this project.

Next, we need the trade flows, $x_{ij}^{k}$ by source country $i$, destination $j$ and industry $k$. We will read the `wiot.fst` file from last class and aggregate the flows.


```{r trade-flows}
library(data.table)
library(fst)

trade_flows <- read_fst("output/wiot.fst", as.data.table = TRUE)
trade_flows <- trade_flows[, by = c("out_ind", "out_country", "in_country"),
                           .(value = sum(value))]
head(trade_flows)
#' Creates the log(1+) of value
trade_flows[, `:=`(
  log1_value = log(1 + value)
  )]
```

_Estimating revealed productivities_ 

We have to estimate the following equation:

$$
\begin{equation}
\ln x_{ij}^k=\delta_{ij}+\delta_{j}^k+\delta_{i}^k+\varepsilon_{ij}^k
\end{equation}
$$

and then invert the fixed effect of exporter-industry to get the productivity $z_i^k$. I extract the revealed productivity of country $i$ in industry $k$ as $z_{i}^k=\exp(\delta_i^k / \theta)$, notice we have normalized all USA productivities and industry _A01	Crop and animal production, hunting and related service activities_ in all other countries to 1, similar to what the original authors have done in Table 2.

Let's create the dummies by ourselves, so they are named when `fixef` is called

```{r}
trade_flows[, `:=`(
  delta_ij = paste0(out_country, "_", in_country),
  delta_jk = paste0(in_country, "_", out_ind),
  delta_ik = paste0(out_country, "_", out_ind)
)]
```

Run the regression of $\log(1+x)$ (But NOT YOU! You should use the PML method to get full grade) in order to have a complete matrix of productivities. Rearrange the USA to be the first country such that the fixed effects are normalized by it. Fixed effects set to zero are the references [(`fixest` documentation)](https://lrberge.github.io/fixest/articles/fixest_walkthrough.html#extracting-the-fixed-effects-coefficients-1), you should check that your fixed-effects are correctly normalized.

```{r}
library(fixest)
eq19 <- feols(
  log1_value~1|delta_ij+delta_jk+delta_ik,
  data = trade_flows[order(-out_country)])
# 
fes <- fixef(eq19)
summary(fes)
```

Now, we extract the revealed productivity of country $i$ in industry $k$ as  $z_{i}^k=\exp(\delta_i^k / \theta)$

```{r zik}
theta <- 6.534
z_ik <- data.table(delta_ik = names(fes$delta_ik), fe = fes$delta_ik)
# Save data.table z_ik for python users
fwrite(z_ik, "output/z_ik.csv")

z_ik[, `:=`(
  z_ik = exp(fe / theta),
  out_country = sub("([[:alpha:]]+)_.*", "\\1", delta_ik),
  out_ind = as.integer(sub("[[:alpha:]]+_(.*)", "\\1", delta_ik)))]
#' Merge back into trade_flows as this is our main dataset. Both z_ik and z_jk
trade_flows <- merge(trade_flows, z_ik[, .(out_country, out_ind, z_ik)],
                     by = c("out_country", "out_ind"),
                     all.x = TRUE)
trade_flows <- merge(trade_flows, z_ik[, .(out_country, out_ind, z_ik)],
                     by.x = c("in_country", "out_ind"),
                     by.y = c("out_country", "out_ind"),
                     all.x = TRUE)
setnames(trade_flows, c("z_ik.x", "z_ik.y"), c("z_ik", "z_jk"))
#' Write trade_flows to fst file
write_fst(trade_flows, "output/trade_flows.fst")
```

## Python code

```{python zik, eval=FALSE}
import numpy as np
import pandas as pd

wiot = pd.read_csv("output/wiot.csv").drop(columns='Unnamed: 0')

wiod_sea = pd.read_excel("input/WIOD_SEA_Nov16.xlsx",
                         sheet_name="DATA",
                         engine="openpyxl")
employed = (wiod_sea
            .query("variable == 'EMP'")
            .groupby('country')[2014]
            .sum()
            .reset_index()
            .rename(columns={'country': 'out_country', 2014: 'employed_i'})
            )
row = pd.DataFrame({'out_country': ['ROW'],
                    'employed_i': [employed['employed_i'].sum()]})
employed = employed.append(row)
employed.to_csv("output/employed.csv", index=False)

# Starting calibration
theta = 6.534

trade_flows = wiot.copy()  # Just to keep names the same as in R code
# Helps in cleaner code
ijk_cols = ['out_country', 'in_country', 'out_ind']
trade_flows = (trade_flows
               .groupby(ijk_cols, as_index=False)[['value']]
               .sum()
               )
trade_flows['log1_value'] = np.log(trade_flows['value'] + 1)
trade_flows['delta_ij'] = trade_flows['out_country'] + \
    '_' + trade_flows['in_country']
trade_flows['delta_jk'] = trade_flows['in_country'] + \
    '_' + trade_flows['out_ind'].astype(str)
trade_flows['delta_ik'] = trade_flows['out_country'] + \
    '_' + trade_flows['out_ind'].astype(str)

# Run a fixed-effect regression to extract the revealed productivities
# This takes 1 minute and crashes
# eq19 = smf.ols('log1_value ~ C(delta_ij)+C(delta_jk)+C(delta_ik)',
#                data=trade_flows).fit()

# # FixedEffect library
# mod = fixedeffect(data_df=trade_flows,
#                   dependent=['log1_value'],
#                   exog_x=[],
#                   category=['delta_ij', 'delta_jk', 'delta_ik'],
#                   cluster=['delta_ij', 'delta_jk', 'delta_ik']).fit()
# mod.summary()
# fes = getfe(mod, normalize=True) # Error

# # pyhdfe library
# ids = trade_flows[['delta_ij', 'delta_jk', 'delta_ik']]
# variables = trade_flows[['log1_value']]
# algorithm = pyhdfe.create(ids)
# residualized = algorithm.residualize(variables)
# residualized
# reg = sm.OLS(residualized, np.ones(len(residualized))).fit(cov_type='HC1')
# reg.summary() # Dead-end, cannot retrieve FEs

# SOLUTION: make the regression in R and load the fixed-effects back
z_ik = pd.read_csv("output/z_ik.csv")
z_ik['z_ik'] = np.exp(z_ik['fe']/theta)
z_ik['out_country'] = z_ik['delta_ik'].str.split('_').apply(lambda x: x[0])
z_ik['out_ind'] = (z_ik['delta_ik'].str.split('_').apply(lambda x: x[1])
                   .astype('int64'))
# Merge back into trade_flows as this is our main dataset. Both z_ik and z_jk
trade_flows = trade_flows.merge(z_ik[['out_country', 'out_ind', 'z_ik']],
                                how='left',
                                left_on=['in_country', 'out_ind'],
                                right_on=['out_country', 'out_ind'])
trade_flows.rename(columns={'z_ik': 'z_jk', 'out_country_x': 'out_country'},
                   inplace=True)
trade_flows = trade_flows.merge(z_ik[['out_country', 'out_ind', 'z_ik']],
                                how='left',
                                on=['out_country', 'out_ind'])
trade_flows.drop('out_country_y', axis=1, inplace=True)
```

## References