---
title: "04 - Trade costs"
author: "Rafael F. Bressan"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    css: "style.css"
bibliography: "references.bib"
link-citations: yes
---

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: { 
      equationNumbers: { 
            autoNumber: "all",
            formatNumber: function (n) {return n}
      } 
  }
});
</script>

```{r sequential, include=FALSE}
targets::tar_read(expenditures)
```

## Calibration - _continued_

```{r loading}
library(data.table)
library(fst)

wiot <- read_fst("output/wiot.fst", as.data.table = TRUE)
trade_flows <- read_fst("output/trade_flows.fst", as.data.table = TRUE)
theta <- 6.543
```

Let's create some auxiliary variables to easily compute trade costs afterwards.
$\delta_{ij}^k \equiv \pi_{ij}^k / \pi_{jj}^k$.
Also, let $\text{z_ratio} = z_i^k / z_j^k$ and $\text{w_ratio} = w_i / w_j$.

```{r}
trade_flows[, `:=`(
  delta_ijk = pi_ijk / pi_jjk,
  z_ratio_ijk = z_ik / z_jk,
  w_ratio_ij = wage_i / wage_j
)]
```

_Computing the trade costs $d_{ij}^k$_

From equation (1) below, valid under assumptions 1 to 4 in @costinot2012goods it's possible to compute all values of $d_{ij}^k$.

$$
\begin{equation}
x_{ij}^k=\frac{(w_id_{ij}^k/z_i^k)^{-\theta}}{\sum_{i'}(w_{i'}d_{i'j}^k/z_{i'}^k)^{-\theta}}\cdot\alpha_j^k w_j L_j
\end{equation}
$$

and given the definition, $\pi_{ij}^k\equiv x_{ij}^k / \sum_i x_{ij}^k$, we arrive at (You should prove it in your notes!):

$$
\begin{equation}
\pi_{ij}^k = \frac{(w_id_{ij}^k/z_i^k)^{-\theta}}{\sum_{i'}(w_{i'}d_{i'j}^k/z_{i'}^k)^{-\theta}}.
\end{equation}
$$

This is a non-linear system of $I\times I\times K$ equations to solve for each $d_{ij}^k$. If we further impose $d_{jj}^k=1$, then it can be easily solved for $d_{ij}^k$ when $i\neq j$ in terms of $\delta_{ij}^k \equiv \pi_{ij}^k / \pi_{jj}^k$, **at the cost of not matching exactly** trade flows $x_{ij}^k$, thus:

$$
\begin{equation}
  \delta_{ij}^k = \left(\frac{d_{ij}^k\cdot w_i/w_j}{z_i^k / z_j^k}\right)^{-\theta}
\end{equation}
$$

and inverting this relation for $d_{ij}^k$ we have:

$$
\begin{equation}
  d_{ij}^k=\frac{\left(\delta_{ij}^k\right)^{-1/\theta}}{w_i/w_j}\cdot z_i^k/z_j^k
\end{equation}
$$

```{r trade-costs}
trade_flows[, d_ijk := ((delta_ijk)^(-1/theta))*z_ratio_ijk/w_ratio_ij]
#' Insert predicted trade flows and shares into trade_flows
trade_flows[, by = c("in_country", "out_ind"),
            Phi_jk := sum((wage_i*d_ijk / z_ik)^(-theta))]
trade_flows[, `:=`(
  pred_value = (((wage_i*d_ijk / z_ik)^(-theta))/Phi_jk)*alpha_jk*wage_j*employed_j,
  pred_pi = ((wage_i*d_ijk / z_ik)^(-theta))/Phi_jk
)]
```


Once we have computed trade costs, we finally have calibrated the model by inputing $\theta$'s value taken from the orignal article to be 6.543, revealed productivities found by means of a regression like equation (6) in [session 2](02_eatonkortum.html), shares $\alpha_j^k$ and $\pi_{ij}^k$ and labor force $L_i$ are calculated (or gathered) from WIOD data, shares of world income and consequently, wages are matched such that [assumption A5](02_eatonkortum.html) holds and the last term computed is trade costs, which is a **residual** term and is found through equation (1) with the additional assumption, $d_{jj}^k=1$. 

Now we can check the calibration quality. We should perform some checks to see if our calculations make sense. For example, $\hat\pi_{ij}^k$ must sum to one over $i$, helding fixed $j$ and $k$. Actually, $\hat\pi_{ij}^k$ must match exactly the ones computed from the data. Moreover, although the predicted trade flows, $\hat x_{ij}^k$, are not guaranteed to match the observed ones, they should at least be highly correlated.  

```{r}
set.seed(657286)
# pi sum to one
trade_flows[, by = c("in_country", "out_ind"), sum(pred_pi)]
# predicted pi equals actual pi
trade_flows[sample(.N, 20), .(out_country, in_country, out_ind, pred_pi, pi_ijk)]
# Correlations
trade_flows[, .(cor(pred_value, value, use = "complete"))]
trade_flows[, .(cor(pred_pi, pi_ijk, use = "complete"))]
```

As you can see above, there are some invalid values for $\hat\pi_{ij}^k$ (i.e. `NaN` or `Inf`). This is due to the fact that some countries do not trade to each other, hence, a few $x_{ij}^{kl}$ trade flows are zero and this may lead to either $\pi_{ij}^k=0$ or $\pi_{jj}^k=0$. **I will leave it to you to figure out how this can be a problem and a way get around it!**

## Python code

```{python tradecosts, eval=FALSE}
# Computing trading costs
trade_flows = (trade_flows
               .assign(delta_ijk=lambda x: x['pi_ijk'] / x['pi_jjk'],
                       z_ratio_ijk=lambda x: x['z_ik'] / x['z_jk'],
                       w_ratio_ij=lambda x: x['wage_i'] / x['wage_j'])
               )
trade_flows = (trade_flows
               .assign(d_ijk=lambda x: x['delta_ijk']**(-1/theta)*x['z_ratio_ijk']/x['w_ratio_ij']))
# Insert predicted trade flows and shares into trade_flows
df = trade_flows[['in_country', 'out_ind', 'wage_i', 'd_ijk', 'z_ik']]
df['Phi_jk'] = (df['wage_i']*df['d_ijk'] / df['z_ik'])**(-theta)
df = df.groupby(['in_country', 'out_ind'], as_index=False)['Phi_jk'].agg(sum)
trade_flows = trade_flows.merge(df, on=['in_country', 'out_ind'], how='left')

# Auxilary functions


def pred_value(df):
    return (((df['wage_i']*df['d_ijk'] / df['z_ik'])**(-theta))/df['Phi_jk'])*df['alpha_jk']*df['wage_j']*df['employed_j']


def pred_pi(df):
    return ((df['wage_i']*df['d_ijk'] / df['z_ik'])**(-theta))/df['Phi_jk']


trade_flows['pred_value'] = pred_value(trade_flows)
trade_flows['pred_pi'] = pred_pi(trade_flows)

# Check pred_pi sums to one
trade_flows.groupby(['in_country', 'out_ind'])['pred_pi'].sum()
# predicted pi equals actual pi
trade_flows[ijk_cols + ['pred_pi', 'pi_ijk']].sample(20)
# Correlations
trade_flows[['pred_pi', 'pi_ijk']].corr()
trade_flows[['pred_value', 'value']].corr()
```

## References